---
title: "Activity 2 - Day 1"
output: github_document
---

```{setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#Install Packages
library(tidyverse)
library(tidymodels)
```

```{r}
#Load data
hfi <- readr::read_csv("https://www.openintro.org/data/csv/hfi.csv")
```

1.  What are the dimensions of the dataset? What does each row
    represent? 
    
**hfi has 1458 rows and 123 columns. Each row represents a country's observation.**
    
```{r}
#Dimensions
dim(hfi)
```

```{r}
hfi_2016 <- hfi %>% filter(year == 2016)
```

2.  What type of plot would you use to display the relationship between
    the personal freedom score, `pf_score`, and `pf_expression_control`?
    
**Since they are continous variables, I would use a scatterplot.**
    
```{r}
#Plot a scatterplot

hfi_2016 %>% 
  ggplot(mapping = aes(y=pf_score, x=pf_expression_control)) +
  geom_point(colour = 'blue', size = 2) +
  labs(
    title = "Relationship between personal freedom score and expression control",
    y = "Personal Freedom Score",
    x = "Personal Freedom Expression Control"
  ) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank()
    )
```

3.  Does the relationship look linear? If you knew a country’s
    `pf_expression_control`, or its score out of 10, with 0 being the
    most, of political pressures and controls on media content, would
    you be comfortable using a linear model to predict the personal
    freedom score?
    
**The relationship looks linear. From the plot it is evident that personal freedom expression control can be used to predict personal freedom score.**

```{r}
#Computes sum of squares
statsr::plot_ss(x = pf_expression_control, y = pf_score, data = hfi_2016, showSquares = TRUE)
```

4.  Using `statsr::plot_ss`, choose a line that does a good job of
    minimizing the sum of squares. Run the function several times. What
    was the smallest sum of squares that you got? How does it compare to
    your neighbour’s? 

**The least SS I got was 3338.187**

``` {r}
m1 <- lm(pf_score ~ pf_expression_control, data = hfi_2016)
tidy(m1)
```
